\texttt{davos}: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2022-10-16 paper/main.tex
fixed incorrect virtual environment citation
WIP implementation details section

At another extreme, virtual machines~\citep{Gold74, AltiEtal05,
  Rose99} provide a hardware-level simulation of the desired system.
Virtual machines are typically isolated from the user's system, such
that installing or running software on a virtual machine does not
impact the user's primary operating system or computing environment.
Containers~\citep[e.g.,][]{Merk14, KurtEtal17} provide a similar
``isolated'' experience.  Although containerized environments do not
specify hardware-level operations, they are typically packaged with a
complete operating system, in addition to a complete copy of Python
and any relevant package dependencies.  Virtual
environments~\citep[e.g.,][]{vanREtal14} also provide a computing
environment that is largely separated from the user's main
environment.  They incorporate a copy of Python and the target
software's dependencies, but virtual environments do not specify or
reproduce an operating system for the runtime environment.  Each of
these systems (virtual machines, containers, and virtual environments)
guarantees (to differing degrees-- at the hardware level, operating
system level, and Python environment level, respectively) that the
relevant code will run similarly for different users.  However, each
of these systems also relies on additional software that can be
resource intensive or burdensome to install or configure.

@@

\subsection{Implementation details}\label{subsec:implementation}
Functionally, importing \texttt{davos} appears to define ``\texttt{smuggle}'' as a Python keyword, similar to ``\texttt{import}'', ``\texttt{def}'', or ``\texttt{return}''. It also appears to cause Python to parse comments as potentially relevant to the code's behavior, which they normally are not. However, \texttt{davos} doesn't actually modify the rules used by Python's parser or lexical analyzer---in fact, modifying the Python grammar isn't possible at runtime, as doing so would require rebuilding the interpreter. Instead, \texttt{davos} leverages the IPython notebook backend to implement the \texttt{smuggle} statement and onion comment via a combination of namespace injections and its own (far simpler) custom parser.

>>>>

At another extreme, virtual machines~\citep{Gold74, AltiEtal05,
  Rose99} provide a hardware-level simulation of the desired system.
Virtual machines are typically isolated from the user's system, such
that installing or running software on a virtual machine does not
impact the user's primary operating system or computing environment.
Containers~\citep[e.g.,][]{Merk14, KurtEtal17} provide a similar
``isolated'' experience.  Although containerized environments do not
specify hardware-level operations, they are typically packaged with a
complete operating system, in addition to a complete copy of Python
and any relevant package dependencies.  Virtual
environments~\citep[e.g.,][]{Anac12} also provide a computing
environment that is largely separated from the user's main
environment.  They incorporate a copy of Python and the target
software's dependencies, but virtual environments do not specify or
reproduce an operating system for the runtime environment.  Each of
these systems (virtual machines, containers, and virtual environments)
guarantees (to differing degrees-- at the hardware level, operating
system level, and Python environment level, respectively) that the
relevant code will run similarly for different users.  However, each
of these systems also relies on additional software that can be
resource intensive or burdensome to install or configure.

@@

\subsection{Implementation details}\label{subsec:implementation}
Functionally, importing \texttt{davos} appears to define ``\texttt{smuggle}'' as a Python keyword, similar to ``\texttt{import}'', ``\texttt{def}'', or ``\texttt{return}''. It also appears to cause comments to be parsed and potentially able to affect the code's behavior, which they normally are not. However, \texttt{davos} doesn't actually modify the rules of Python's parser or lexical analyzer---in fact, modifying the Python grammar isn't possible at runtime, as doing so would require rebuilding the interpreter. Instead, \texttt{davos} leverages the IPython notebook backend to implement the \texttt{smuggle} statement and onion comment via a combination of namespace injections and its own (far simpler) custom parser.

%The \texttt{smuggle} keyword can be enabled and disabled at will by ``activating'' and ``deactivating'' \texttt{davos} (see Section \ref{subsec:config}, above). When \texttt{davos} is imported, it is automatically activated by default. Activating \texttt{davos} triggers two things: 
%\begin{enumerate}
%    \item The \texttt{smuggle()} function is injected into the IPython user namespace 
%    \item The \texttt{davos} parser is registered as a custom IPython input transformer
%\end{enumerate}

The \texttt{smuggle} keyword can be enabled and disabled at any time by ``activating'' and ``deactivating'' \texttt{davos} (see Section \ref{subsec:config}, above); when \texttt{davos} is first imported, it is activated automatically. Activating \texttt{davos} triggers two things: (1) the \texttt{smuggle()} function is injected into the IPython user namespace, and (2) the \texttt{davos} parser is registered as a custom IPython input transformer. IPython preprocesses all executed code as plain text before it is sent to the Python parser, in order to handle special constructs like \texttt{\%magic} and \texttt{!shell} commands. \texttt{davos} hooks into this process to transform \texttt{smuggle} statements into syntactically valid Python code. The \texttt{davos} parser uses a complex regular expression \cite{Thom68} to match lines of code containing \texttt{smuggle} statements (and, optionally, onion comments), extract relevant information from their text, and replace them with equivalent calls to the \texttt{smuggle()} function. For example, if the code visible to the user reads 
\begin{verbatim}
smuggle numpy as np    # pip: numpy>1.16,<=1.20 -vv,
\end{verbatim} 
the code that is actually executed by the Python interpreter would be 
\begin{verbatim}
smuggle(name="numpy", as_="np", installer="pip", args_str="""numpy>1.16,<=1.20 -vv""", installer_kwargs={`editable': False, `spec': `numpy>1.16,<=1.20', `verbosity': 2}).
\end{verbatim}

% talk about how exceptions have to inherit from SyntaxError since that's the only thing that can be raised during parsing
